\Title

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

We would like to highlight a few innovations in the contributions presented in this paper. First, while~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatiotemporal motifs. Moreover, compared to a representation using this algorithm, the parameters of the model (weights and delays) are explicable, as they directly inform about the logit (inverse sigmoid of the probability) of detection for each spatiotemporal spike motif. Another novelty is that the model learns the weights and delays simultaneously. For example, the polychronization model~\citep{izhikevich_polychronization_2006} learns only the weights using STDP, while the delays are randomly drawn and their values are frozen during learning. In addition, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified toy problems~\citep{gutig_tempotron_2006}. Another major contribution is to provide a model that is suitable for learning any kind of spatiotemporal spiking motif and that can be trained in a supervised manner by providing a dataset of supervised pairs. Instead of relying on a careful description of the physical rules governing a task, e.g. the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}, this allows a more flexible definition of the model using this properly labeled dataset.
%
\subsection{Main limits}
%%%-----------------------------------------------------------------
We have identified a number of limitations of our model, which we will now discuss in detail. First, the entire framework is based on a discrete binning of time, which is not compatible with the continuous nature of biological time. We used this binning to efficiently implement the framework on conventional hardware, especially GPUs, to be able to use fast three-dimensional convolutions. We have tested the effect of the width of the time bin and observed it has essentially no effect on the results presented in this paper. This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where accuracy was unaffected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}. This suggests the possibility of analytically including a precision term in the temporal value of the input spikes. This mechanism may be implemented by the filtering implemented by the synaptic time constant of about $5~\ms$. Furthermore, it is possible to circumvent the need for time discretization by the use of a purely event-based scheme. In fact, it is not necessary to compute the voltage traces between two spikes~\citep{hanuschkin_general_2010}. Thus, it is possible to define a purely event-based framework. Such an architecture could provide promising computational speedups.

>>>>

We would like to highlight a few innovations in the contributions presented in this paper. First, while~\citep{ghosh_spatiotemporal_2019,yu_stsc-snn_2022} use a correlation-based heuristic, the generic heterogeneous model is formalized from first principles for optimal detection of the event-based spatiotemporal motifs. Moreover, compared to a representation using this algorithm, the parameters of the model (weights and delays) are explicable, as they directly inform about the evidence  of detection for each spatiotemporal spike motif. This evidence is measured by the logit of the probability, that is, the inverse sigmoid of the probability. Another novelty is that the model learns the weights and delays simultaneously. For example, the polychronization model~\citep{izhikevich_polychronization_2006} learns only the weights using STDP, while the delays are randomly drawn and their values are frozen during learning. In addition, the model is evaluated on a realistic task, while models such as the tempotron are tested on simplified toy problems~\citep{gutig_tempotron_2006}. Another major contribution is to provide a model that is suitable for learning any kind of spatiotemporal spiking motif and that can be trained in a supervised manner by providing a dataset of supervised pairs. Instead of relying on a careful description of the physical rules governing a task, e.g. the luminance conservation principle for motion detection~\citep{benosman_asynchronous_2012, dardelet_event-by-event_2021}, this allows a more flexible definition of the model using this properly labeled dataset.
%
\subsection{Main limits}
%%%-----------------------------------------------------------------
We have identified a number of limitations of our model, which we will now discuss in detail. First, the entire framework is based on a discrete binning of time, which is not compatible with the continuous nature of biological time. We used this binning to efficiently implement the framework on conventional hardware, especially GPUs, to be able to use fast three-dimensional convolutions. We have tested the effect of the width of the time bin and observed it has essentially no effect on the results presented in this paper. This is consistent with the relative robustness of other event-based frameworks such as HOTS~\citep{lagorce_hots_2017}, where accuracy was unaffected when the input spikes were subjected to noisy perturbations up to $1~\ms$~\citep{grimaldi_robust_2022}. This suggests the possibility of analytically including a precision term in the temporal value of the input spikes. This mechanism may be implemented by the filtering implemented by the synaptic time constant of about $5~\ms$. Furthermore, it is possible to circumvent the need for time discretization by the use of a purely event-based scheme. In fact, it is not necessary to compute the voltage traces between two spikes~\citep{hanuschkin_general_2010}. Thus, it is possible to define a purely event-based framework. Such an architecture could provide promising computational speedups.
